Sequoidal Categories and Transfinite Games: Towards a Coalgebraic Approach to Linear Logic\footnote{This work was partially supported by UK EPSRC Grant EP/K037633/1}

### 2017-04-11 round-corner/paper/lipics-v2016-authors/sequoidal.tex
Fixed up a few things.

% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Sequoidal Categories and Transfinite Games: Towards a Coalgebraic Approach to Linear Logic\footnote{This work was partially supported by [TODO: find the appropriate grant codes]}}
\titlerunning{Sequoidal Categories and Transfinite Games} %optional, in case that the title is too long; the running title should fit into the top page column

@@

For example, consider the category $\Rel$ of sets and relations, with symmetric monoidal structure given by the cartesian product (with unit $I$, the singleton set $\{*\}$). This has finite (bi)products (disjoint unions) so we may define the functor $F(A,S) = (A \otimes S) \oplus I$. For any object (set) $A$, let $A^*$ be the set of finite sequences of elements of $A$ (i.e. the carrier of the free monoid on $A$), and $\alpha:A^* \rightarrow F(A,A^*)$ be the morphism $\{(\varepsilon,\inr(*)\}\cup  \{(aw,(\inl(a,w))\ |\ a \in A,w \in A^*\}$. It is straightforward to show that: 
% $A^*$ is a \emph{minimal invariant} for $F(A,\_)$ with respect to the inclusion order on $\Rel$ and so:
\begin{lemma}$(A^*,\alpha_A)$ is the final coalgebra  for $F(A,\_)$.
\end{lemma}
Since we have a natural transformation  $\inl_{A,S}:A \otimes S \rightarrow F(A,S)$, we may encapsulate the state in the  state transformer $f:S \rightarrow A \otimes S$ by taking the \emph{anamorphism} of $f;\inl_{A,B}:S \rightarrow F(A,S)$, --- i.e. the unique $F(A,\_)$-coalgebra morphism from $(S,f;\inl)$ into $(A^*,\alpha_{A})$. This  is  a morphism from an initial state $S$ into $A^*$: by definition, composing it with $\alpha:A \rightarrow F(A,A^*)$ (which we can think of as  \emph{invoking} our stateful object) returns a copy of $f$ and together with the encapsulated morphism with updated internal state.

%By distributivity of $\oplus$ over  $\otimes$ we have a natural transformation $t_{A,S}: F(A,S) \otimes S' \rightarrow F(A,S \otimes S')$. We may use this to encapsulate  state transformers of general form: given $f:A \otimes S \rightarrow B \otimes S$, taking the anamorphism of the $F(B,\_)$-colagebra:    
%$$A^* \otimes S  \stackrel{\alpha_A \otimes S}{\longrightarrow} F(A,A^*) \otimes S \stackrel{t}{\longrightarrow} F(A,A^* \otimes S)$$

 Distributivity of $\oplus$ over $\otimes$ implies that $F(A \oplus A',S) \cong F(A,S) \oplus F(A',S)$.  This allows state transformers to be aggregated, to construct stateful objects compounded of a series of methods which share access to a common state. 
%i.e. given morphisms $f_1:A_1 \otimes S \rightarrow B_1 \otimes S, \ldots , f_n:A_n \otimes S \rightarrow B_n \otimes S$, encapsulating  $\dist;(f_1 \oplus \ldots \oplus f_n);\dist^{-1}$, gives a morphism from $(A_1 \oplus \ldots \oplus A_n)^* \otimes S$ into $(B_1 \oplus \ldots \oplus B_n)^*$.          
For example, we may represent a reference cell storing integer values as a state transformer $\cell:\Na \rightarrow (\Na \oplus \Na) \otimes \Na$, obtained by aggregating two ``methods'' which share access to a value in $\Na$ representing the contensts of the cell --- returning a ``read'' of the input state (and leaving it unchanged) or accepting a ``write'' of a new value and using it to update the state.   Thus (with appropriate tagging) it is the relation $\{(i, (\read(i),i))\ |\ i \in \Na\} \cup \{(i,{\mathtt{write}}(j),j)\}$.
The anamorphism of the coalgebra $\cell;\inl:\Na \rightarrow F(\Na \oplus \Na,\Na) $  is the relation from $\Na$ to $(\Na \oplus \Na)^*$ consisting of pairs of the form $(i_1,\read(i_1)^*{\mathtt{write}}(i_2)\read(i_2)^*\ldots)$. Composition with this morphism is precisely the interpretation of new variable declaration in the semantics  in $\Rel$ of the prototypical functional-imperative language \emph{Syntactic Control of Interference} (SCI) given in \cite{Mcsci}.

@@

%However,  the richer structure of
%game semantics does support a coalgebraic cofree commutative comonoid. 
% As we shall show by example and counterexample, the richer structure the relationship between initial and final coalgebras, and cofree objects is more subtle.
%  (note, for example, that any final coalgebra in $\Rel$ must be bifree, by its self-duality).

@@

We now outline the remainder of the paper. Our main contribution is an investigation of the circumstances in which the  cofree commutative comonoid on $A$ arises from a final coalgebra for the functor $A\sequoid\blank$, where $\sequoid$ (the sequoid) is a non-commutative operation on games  introduced by one of the authors \cite{laird02}.\footnote{We will focus on a particular category of ``history sensitive'', Abramsky-Jagadeesan style games \cite{abramskyjagadeesangames}, but sequoidal structure is a unifying feature of sequential, history-sensitive games.} %Under a simple interpretation, $A\sequoid B$ is the same as the game $A\tensor B$, except that the first move must take place in $A$. 

@@

\subsection{Game semantics and the sequoidal operator}

To get around the problems caused by the non-commutativity of the comonoid in the sets-and-relations model, we shall consider a game semantics model, in which we will be able to construct cofree commutative comonoids.  We shall present a form of game semantics in the style of \cite{hyland1997games} and \cite{abramskyjagadeesangames}.  A game $A$ is given by a forest (i.e., a prefix-closed set) $P_A$ of \emph{positions}, which are alternating sequences of $O$-moves and $P$-moves.  We call a position a \emph{$P$-position} if it ends with a $P$-move and an \emph{$O$-position} if it ends with an $O$-move.  A \emph{strategy} for a game $A$ is a subforest $\sigma$ of $P_A$ that is closed under $O$-replies to $P$-positions and which satisfies \emph{determinism}: if $sa,sb\in\sigma$, where $s$ is a $P$-position, then $a=b$.  We will adopt the rule that all positions must start with an $O$-move.

@@

\begin{definition}
  Let $\C=(\C,\C_s,J,\wk)$ be a sequoidal category.  We say that $\C$ is an \emph{inclusive sequoidal category} if $\C_s$ is a full-on-objects subcategoryof $\C$ containing all isomorphisms in $\C$ and the morphisms $\wk_{A,B}$ and $J$ is the inclusion functor.

  If $\C$ is an inclusive sequoidal category, we say that $\C$ is \emph{Cartesian} if $\C_s$ has all products and these are preserved by $J$.  In that case, we say that $\C$ is \emph{decomposable} if the natural transformations
  \begin{gather*}
    \dec_{A,B} = \langle \wk_{A,B}, \comp{\wk_{A,B}}{\sym_{A,B}}\rangle\from A\tensor B\to (A\sequoid B)\times (B\sequoid A) \\
    \dec^0 \from I \to 1
  \end{gather*}
  are isomorphisms and we say that $\C$ is \emph{distributive} if the natural transformations
  \begin{gather*}
    \dist_{A,B,C} = \langle \pr_1\sequoid \id_C,\pr_2\sequoid\id_C\rangle\from (A\times B)\sequoid C\to (A\sequoid C)\times (B\sequoid C) \\
    \dist_{(A_i\colon i\in I),C} = \langle (\pr_i\sequoid\id_C\suchthat i\in I)\rangle \from \left(\prod_{i\in I} A_i\right) \sequoid C \to \prod_{i\in I}(A_i\sequoid C) \\
    \dist_{A,0}\from 1\sequoid A\to 1
  \end{gather*}
  are isomorphisms.
\end{definition}
\begin{definition}
  Let $\C=(\C,\C_s,J,\wk)$ be a sequoidal category.  We say that $\C$ is an \emph{inclusive sequoidal category} if $\C_s$ is a full-on-objects subcategory of $\C$ containing all isomorphisms and finite products of $\C$, and the morphisms $\wk_{A,B}$ and $J$ is the inclusion functor.

 We say that $\C$ is \emph{decomposable} if for any $A$ and $B$, the tensor product $A \tensor B$ is a cartesian product of $A \sequoid B$ and $B \sequoid A$, with projections $\wk_{A,B}:A \tensor B \rightarrow A \sequoid B$ and  $\comp{\wk_{A,B}}{\sym_{A,B}}:A \tensor B \rightarrow B \sequoid A$, and if the product $A \times B$ exists, then $(A \times B) \sequoid C$ is the product of $A \sequoid C$ and $B \sequoid C$, with projections $\pr_l \sequoid \id_C,\pr_r \sequoid \id_C$. 
So if $\C$ has all finite products, 
the natural transformations
  \begin{gather*}
    \dec_{A,B} = \langle \wk_{A,B}, \comp{\wk_{A,B}}{\sym_{A,B}}\rangle\from A\tensor B\to (A\sequoid B)\times (B\sequoid A) \\
   % \dec^0 \from I \to 1 \\
    \dist_{A,B,C} = \langle \pr_1\sequoid \id_C,\pr_2\sequoid\id_C\rangle\from (A\times B)\sequoid C\to (A\sequoid C)\times (B\sequoid C) \\
    \dist_{A,0}\from 1\sequoid A\to 1
  \end{gather*}
  are isomorphisms.
\end{definition}

We shall also use the fact that our category of games is \emph{affine} - so the natural transformation $\dec^0\from I \to 1$ is an isomorphism.  This is not strictly necessary, and techniques are available to obtain similar results in the non-affine case (see \cite{jimNonAffine}, for example), but it will greatly simplify our presentation.

The games model satisfies an analog of monoidal closedness, but we will not give that here (see \cite{laird02}).  We see that $\G$ is a prototype example of a distributive, decomposable sequoidal category.

@@

In the category of games, the morphisms $\id\sequoid\pr_1$ and $\id\sequoid\pr_2 \from A \sequoid (B\times C)\to (A \sequoid B) \times (A \sequoid C)$ are jointly monomorphic.  If a distributive sequoidal category satisfies the same property, we say that it is \emph{strong distributive}.
\begin{proposition}Any strong distributive decomposabe sequoidal category has all symmetric tensor powers, and these are preserved by the tensor.  
\end{proposition} 
\begin{proof}
By sequoidal decomposability, for any $n$, $A^{\tensor (n+1)}$ is the cartesian product $\Pi_{i \le n} (A \sequoid A^{\tensor n})$ with projections 
$\sym_i;\wk_{A,A^{\tensor n}} $, where  $\sym_i: A^{\tensor (n+1)} \rightarrow A^{\tensor (n+1)}$ is the symmetry isomorphism corresponding to the permutation on $n$ which swaps $1$ and $i$.

@@

We define a \emph{winning strategy} on $(A,\zeta_A)$ to be a total strategy $\sigma$ on $A$ such that $\zeta_A(s)=P$ whenever $s$ is an infinite sequence, all of whose finite prefixes are in $\sigma$.  It is known (see \cite{abramskyjagadeesangames}) that the composition of winning strategies is winning and that we get a decomposable, distributive sequoidal closed category $\W$ with $\oc A$ as the final coalgebra for $A\sequoid\blank$ and the cofree commutative comonoid over $A$ \cite{martinsthesis}.

@@

\subsection{The Sequoidal Exponential as a Bifree Algebra}
Observe that in our category of games, $(!A,\alpha)$ is in fact a \emph{bifree algebra} for $A\sequoid\blank$ --- the isomorphism $\alpha^{-1}: A \sequoid !A \rightarrow !A$ is an initial algebra for  $A \sequoid \blank$.  We may show that in such cases, the condition that $!$ is strong monoidal --- and thus the cofree exponential ---  always holds\footnote{Without requiring our sequoidally decomposable category to have finite products  we may equip each object $\oc A$ with the structure of a comonoid by defining:   
$\mu:!A \rightarrow !A  \otimes !A$
 to be the catamorphism of the $A \sequoid \blank$ algebra:
\[
  A \sequoid (!A \otimes !A) %\stackrel{\langle \id, A \sequoid \sym\rangle}{\cong} 
\cong A \sequoid (!A \otimes !A) \times  A \sequoid (!A \otimes !A) \cong (A \sequoid !A) \sequoid !A \times  (A \sequoid !A) \sequoid !A \cong (!A \sequoid !A) \cong (!A \sequoid !A) \cong (!A \otimes !A)
\]
This satisfies the further requirements of a \emph{linear category} in the sense of \cite{Bierman}, although it does not appear to be possible to show that it is the cofree commutative comonoid.} :
we may define an inverse to $\int:!A \otimes !A \rightarrow ! (A \times B)$  as the \emph{catamorphism} of the  $A \sequoid\blank$-algebra:
$$(A \times B) \sequoid !A \otimes !B \cong (A \sequoid (!A \otimes !B)) \times  (B \sequoid (!A \otimes !B)) \cong (!A \sequoid !B) \times (!B \sequoid !A)$$

It is not necessary for the final $A \sequoid\_$-coalgebra to be bifree for the exponential to be strong monoidal and thus the cofree commutative comonoid. An example is provided by  the category $\W$ of win-games and winning strategies, which is sequoidal closed and decomposable. To show that the  final $A \sequoid\_$-coalgebra in this category is not bifree, it suffices to observe that from such an algebra, we may derive a  \emph{fixed point} operator $\fix_A:\C(A,A) \rightarrow \C(I,A)$ for each $A$, such that $\fix_A(f);f = \fix(f)$.      
\begin{proposition}Suppose $\C$ is sequoidal closed and decomposable, and $(!A,\alpha)$ is a bifree $A \sequoid\_$-algebra. Then we may define a fixed point operator on $\C$.  
\end{proposition}    
\begin{proof}For any $A$, let  $\Phi_A:!(A \multimap A) \rightarrow A$ be the catamorphism of the counit to the adjunction $A \sequoid\blank \dashv A \multimap$,  $\epsilon_{A,A}: (A \multimap A) \sequoid A \rightarrow A$, which is a  $(A \multimap A) \sequoid \blank$-algebra. For any morphism $f:A \rightarrow A$ we may define $\fix_A(f) = \Lambda(f)^\dag;\Phi_A$, where $\Lambda(f):I \rightarrow (A \multimap A)$ is the ``name'' of $f$.    
\end{proof}
As one would expect, it is not possible to define a fixed point operator on the category of games and winning strategies ---  for example, if $\bot$ is the game with a single move then the hom-set $\C(I,\bot)$ is empty and hence there can be no morphism $\fix_\bot(\id_\bot)$. So the final  $A \sequoid \_$-coalgebra is not bifree in this case.

@@

In particular, $\oc A$ is the final coalgebra for $A\sequoid\blank$.  But, as we showed above, it is not the cofree commutative coalgebra over $A$.

@@

By Lemma \ref{aFormulaForAlpha}, the morphism along the bottom is equal to $\alpha_A$ and therefore $g$ is the anamorphism for the morphism along the top; i.e., $g=f^\dag$.  
\end{document}

>>>>

% Author macros::begin %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title{Sequoidal Categories and Transfinite Games: Towards a Coalgebraic Approach to Linear Logic\footnote{This work was partially supported by UK EPSRC Grant EP/K037633/1}}
\titlerunning{Sequoidal Categories and Transfinite Games} %optional, in case that the title is too long; the running title should fit into the top page column

@@

For example, consider the category $\Rel$ of sets and relations, with symmetric monoidal structure given by the cartesian product (with unit $I$, the singleton set $\{*\}$). This has finite (bi)products (disjoint unions) so we may define the functor $F(A,S) = (A \otimes S) \oplus I$. For any object (set) $A$, let $A^*$ be the set of finite sequences of elements of $A$ (i.e. the carrier of the free monoid on $A$), and $\alpha:A^* \rightarrow F(A,A^*)$ be the morphism $\{(\varepsilon,\inr(*)\}\cup  \{(aw,(\inl(a,w))\ |\ a \in A,w \in A^*\}$. It is straightforward to show that: 
\begin{lemma}$(A^*,\alpha_A)$ is the final coalgebra  for $F(A,\_)$.
\end{lemma}
Since we have a natural transformation  $\inl_{A,S}:A \otimes S \rightarrow F(A,S)$, we may encapsulate the state in the  state transformer $f:S \rightarrow A \otimes S$ by taking the \emph{anamorphism} of $f;\inl_{A,B}:S \rightarrow F(A,S)$, --- i.e. the unique $F(A,\_)$-coalgebra morphism from $(S,f;\inl)$ into $(A^*,\alpha_{A})$. This  is  a morphism from an initial state $S$ into $A^*$: by definition, composing it with $\alpha:A \rightarrow F(A,A^*)$ (which we can think of as  \emph{invoking} our stateful object) returns a copy of $f$ and together with the encapsulated morphism with updated internal state.

 Distributivity of $\oplus$ over $\otimes$ implies that $F(A \oplus A',S) \cong F(A,S) \oplus F(A',S)$.  This allows state transformers to be aggregated, to construct stateful objects compounded of a series of methods which share access to a common state. 
For example, we may represent a reference cell storing integer values as a state transformer $\cell:\Na \rightarrow (\Na \oplus \Na) \otimes \Na$, obtained by aggregating two ``methods'' which share access to a value in $\Na$ representing the contensts of the cell --- returning a ``read'' of the input state (and leaving it unchanged) or accepting a ``write'' of a new value and using it to update the state.   Thus (with appropriate tagging) it is the relation $\{(i, (\read(i),i))\ |\ i \in \Na\} \cup \{(i,{\mathtt{write}}(j),j)\}$.
The anamorphism of the coalgebra $\cell;\inl:\Na \rightarrow F(\Na \oplus \Na,\Na) $  is the relation from $\Na$ to $(\Na \oplus \Na)^*$ consisting of pairs of the form $(i_1,\read(i_1)^*{\mathtt{write}}(i_2)\read(i_2)^*\ldots)$. Composition with this morphism is precisely the interpretation of new variable declaration in the semantics  in $\Rel$ of the prototypical functional-imperative language \emph{Syntactic Control of Interference} (SCI) given in \cite{Mcsci}.

@@

@@

We now outline the remainder of the paper. Our main contribution is an investigation of the circumstances in which the  cofree commutative comonoid on $A$ arises from a final coalgebra for the functor $A\sequoid\blank$, where $\sequoid$ (the sequoid) is a non-commutative operation on games  introduced by one of the authors \cite{laird02}.\footnote{We will focus on a particular category of ``history sensitive'', Abramsky-Jagadeesan style games \cite{abramskyjagadeesangames}, but sequoidal structure is a unifying feature of sequential, history-sensitive games.} 

@@

\subsection{Game semantics and the sequoidal operator}
We shall present a form of game semantics in the style of \cite{hyland1997games} and \cite{abramskyjagadeesangames}.  A game $A$ is given by a forest (i.e., a prefix-closed set) $P_A$ of \emph{positions}, which are alternating sequences of $O$-moves and $P$-moves.  We call a position a \emph{$P$-position} if it ends with a $P$-move and an \emph{$O$-position} if it ends with an $O$-move.  A \emph{strategy} for a game $A$ is a subforest $\sigma$ of $P_A$ that is closed under $O$-replies to $P$-positions and which satisfies \emph{determinism}: if $sa,sb\in\sigma$, where $s$ is a $P$-position, then $a=b$.  We will adopt the rule that all positions must start with an $O$-move.

@@

\begin{definition}
  Let $\C=(\C,\C_s,J,\wk)$ be a sequoidal category.  We say that $\C$ is \emph{inclusive} if $\C_s$ is a full-on-objects subcategoryof $\C$ containing all isomorphisms in $\C$ and the morphisms $\wk_{A,B}$, and $J$ is the inclusion functor.

 We say that $\C$ is \emph{decomposable} if for any $A$ and $B$, the tensor product $A \otimes B$ is a cartesian product of $A \sequoid B$ and $B \sequoid A$, with projections $\wk_{A,B}:A \otimes B \rightarrow A \sequoid B$ and  $\comp{\wk_{A,B}}{\sym_{A,B}}:A \otimes B \rightarrow B \sequoid A$. $\C$ is \emph{distributive}  if whenever the product $A \times B$ exists, then $(A \times B) \sequoid C$ is the product of $A \sequoid C$ and $B \sequoid C$, with projections $\pi_l \sequoid C,\pi_r \sequoid C$. 
So if $\C$ is Cartesian (has all finite products), 
the natural transformations
  \begin{gather*}
    \dec_{A,B} = \langle \wk_{A,B}, \comp{\wk_{A,B}}{\sym_{A,B}}\rangle\from A\tensor B\to (A\sequoid B)\times (B\sequoid A) \\
    \dist_{A,B,C} = \langle \pr_1\sequoid \id_C,\pr_2\sequoid\id_C\rangle\from (A\times B)\sequoid C\to (A\sequoid C)\times (B\sequoid C) \\
    \dist_{A,0}\from 1\sequoid A\to 1
  \end{gather*}
  are isomorphisms.
\end{definition}

\begin{definition}
  Let $\C=(\C,\C_s,J,\wk)$ be a sequoidal category.  We say that $\C$ is an \emph{inclusive sequoidal category} if $\C_s$ is a full-on-objects subcategory of $\C$ containing all isomorphisms and finite products of $\C$, and the morphisms $\wk_{A,B}$ and $J$ is the inclusion functor.

 We say that $\C$ is \emph{decomposable} if $I$ is a terminal object for $\C$ and if for any $A$ and $B$, the tensor product $A \tensor B$ is a cartesian product of $A \sequoid B$ and $B \sequoid A$, with projections $\wk_{A,B}:A \tensor B \rightarrow A \sequoid B$ and  $\comp{\wk_{A,B}}{\sym_{A,B}}:A \tensor B \rightarrow B \sequoid A$, and if the product $A \times B$ exists, then $(A \times B) \sequoid C$ is the product of $A \sequoid C$ and $B \sequoid C$, with projections $\pr_l \sequoid \id_C,\pr_r \sequoid \id_C$. 
So if $\C$ has all finite products, 
the natural transformations
  \begin{gather*}
    \dec_{A,B} = \langle \wk_{A,B}, \comp{\wk_{A,B}}{\sym_{A,B}}\rangle\from A\tensor B\to (A\sequoid B)\times (B\sequoid A) \\
    \dec^0 \from I \to 1 \\
    \dist_{A,B,C} = \langle \pr_1\sequoid \id_C,\pr_2\sequoid\id_C\rangle\from (A\times B)\sequoid C\to (A\sequoid C)\times (B\sequoid C) \\
    \dist_{A,0}\from 1\sequoid A\to 1
  \end{gather*}
  are isomorphisms.
\end{definition}

We shall also use the fact that our category of games is \emph{affine} - so the natural transformation $\dec^0\from I \to 1$ is an isomorphism.  This is not strictly necessary, and techniques are available to obtain similar results in the non-affine case (see \cite{jimNonAffine}, for example), but it will greatly simplify our presentation.
 We see that $\G$ is a prototype example of a distributive, decomposable sequoidal category.

@@

In the category of games, the morphisms $\id\sequoid\pr_1$ and $\id\sequoid\pr_2 \from A \sequoid (B\times C)\to (A \sequoid B) \times (A \sequoid C)$ are jointly monomorphic.  If a distributive sequoidal category satisfies the same property, we say that it is \emph{strong distributive}.
\begin{proposition}Any strong distributive decomposable sequoidal category has all symmetric tensor powers, and these are preserved by the tensor.  
\end{proposition} 
\begin{proof}
By sequoidal decomposability, for any $n$, $A^{\tensor (n+1)}$ is the cartesian product $\Pi_{i \le n} (A \sequoid A^{\tensor n})$ with projections 
$\sym_i;\wk_{A,A^{\tensor n}} $, where  $\sym_i: A^{\tensor (n+1)} \rightarrow A^{\tensor (n+1)}$ is the symmetry isomorphism corresponding to the permutation on $n$ which swaps $1$ and $i$.

@@

We define a \emph{winning strategy} on $(A,\zeta_A)$ to be a total strategy $\sigma$ on $A$ such that $\zeta_A(s)=P$ whenever $s$ is an infinite sequence, all of whose finite prefixes are in $\sigma$.  It is known (see \cite{abramskyjagadeesangames}) that the composition of winning strategies is winning and that we get a decomposable, distributive sequoidal category $\W$ with $\oc A$ as the final coalgebra for $A\sequoid\blank$ and the cofree commutative comonoid over $A$ \cite{martinsthesis}.

@@

\subsection{The Sequoidal Exponential as a Bifree Algebra}
Observe that in our category of games, $(!A,\alpha)$ is in fact a \emph{bifree algebra} for $A\sequoid\blank$ --- the isomorphism $\alpha^{-1}: A \sequoid !A \rightarrow !A$ is an initial algebra for  $A \sequoid \blank$.  We may show that in such cases, the condition that $!$ is strong monoidal --- and thus the cofree exponential ---  always holds\footnote{Without requiring our sequoidally decomposable category to have finite products  we may equip each object $\oc A$ with the structure of a comonoid by defining:   
$\mu:!A \rightarrow !A  \otimes !A$
 to be the catamorphism of the $A \sequoid \blank$ algebra:
\[
  A \sequoid (!A \otimes !A) 
\cong A \sequoid (!A \otimes !A) \times  A \sequoid (!A \otimes !A) \cong (A \sequoid !A) \sequoid !A \times  (A \sequoid !A) \sequoid !A \cong (!A \sequoid !A) \cong (!A \sequoid !A) \cong (!A \otimes !A)
\]
This satisfies the further requirements of a \emph{linear category} in the sense of \cite{Bierman}, although it does not appear to be possible to show that it is the cofree commutative comonoid.} :
we may define an inverse to $\int:!A \otimes !A \rightarrow ! (A \times B)$  as the \emph{catamorphism} of the  $A \sequoid\blank$-algebra:
$$(A \times B) \sequoid !A \otimes !B \cong (A \sequoid (!A \otimes !B)) \times  (B \sequoid (!A \otimes !B)) \cong (!A \sequoid !B) \times (!B \sequoid !A)$$

It is not necessary for the final $A \sequoid\_$-coalgebra to be bifree for the exponential to be strong monoidal and thus the cofree commutative comonoid. An example is provided by  the category $\W$ of win-games and winning strategies, which is \emph{sequoidal closed} (the restriction of the functor  $A \multimap\blank$ to strict strategies is right adjoint to $\blank \sequoid A$, and inclusion sends this adjunction to the usual monoidal closure). To show that the  final $A \sequoid\blank$-coalgebra in this category is not bifree, it suffices to observe that from such an algebra, we may derive a  \emph{fixed point} operator $\fix_A:\C(A,A) \rightarrow \C(I,A)$ for each $A$, such that $\comp{f}{\fix_A(f)} = \fix(f)$.      
\begin{proposition}Suppose $\C$ is sequoidal closed and decomposable, and $(!A,\alpha)$ is a bifree $A \sequoid\_$-algebra. Then we may define a fixed point operator on $\C$.  
\end{proposition}    
\begin{proof}For any $A$, let  $\Phi_A:!(A \multimap A) \rightarrow A$ be the catamorphism of the counit to the adjunction $A \sequoid\blank \dashv A \multimap$,  $\epsilon_{A,A}: (A \multimap A) \sequoid A \rightarrow A$, which is a  $(A \multimap A) \sequoid \blank$-algebra. For any morphism $f:A \rightarrow A$ we may define $\fix_A(f) = \Lambda(f)^\dag;\Phi_A$, where $\Lambda(f):I \rightarrow (A \multimap A)$ is the ``name'' of $f$.    
\end{proof}
As one would expect, it is not possible to define a fixed point operator on the category of games and winning strategies ---  for example, if $\bot$ is the game with a single move then the hom-set $\C(I,\bot)$ is empty and hence there can be no morphism $\fix_\bot(\id_\bot)$. So the final  $A \sequoid \_$-coalgebra is not bifree in this case.

@@

In particular, $\oc A$ is the carrier of the final coalgebra for $A\sequoid\blank$.  But, as we saw before, it is not the carrier for the cofree commutative comonoid over $A$.

@@

By Lemma \ref{aFormulaForAlpha}, the morphism along the bottom is equal to $\alpha_A$ and therefore $g$ is the anamorphism for the morphism along the top; i.e., $g=f^\dag$.

\end{document}
